Abstractions for Multi-Sorted Substitutions

### 2023-09-20 paper/paper.tex
Update.

  To fill this gap, we introduce a new abstraction that we call a
  \ACode{\ACon{ComposeKit}},
  which provides an operation \ACode{\AField{\_\&/⋯\_}} that unifies
  \ACode{\ADef{\_\&\_}} and \ACode{\ADef{\_⋯\_}}. This allows us to
  define a general composition, which informally can be described as:
  \ExampleCode\KComposition
  We then define a \ACode{\ACon{ComposeTraversal}}, which consists of the
  \ACode{\AField{⋯-fusion}} lemma for general maps.
  Just like a \ACode{\ACon{Traversal}} allowed us to bootstrap the two
  \ACode{\ACon{Kit}} instances, the \ACode{\ACon{ComposeTraversal}} allows
  us to bootstrap the four \ACode{\ACon{ComposeKit}} instances.

@@

  The \ACode{\ACon{ComposeKit}} abstraction provides the
  \ACode{\AField{\_\&/⋯\_}} operation, necessary for defining
  composition, and two lemmas about its behavior:
  \LibCode\KComposeKit
  The \ACode{\AField{\_\&/⋯\_}} operation takes a variable or term
  (according to \ACode{K₁}) and a renaming or substitution
  (according to \ACode{K₂}) and performs the map traversal returning
  a variable or term (according to \ACode{K₁⊔K₂}).
  We later define four instances for \ACode{\ACon{ComposeKit}}, where
  \ACode{K₁} and \ACode{K₂} can each be either the variable or term kit.
  The third parameter \ACode{K₁⊔K₂} can be seen as a functional
  dependency and is completely determined by the choice of \ACode{K₁}
  and \ACode{K₂}.

@@

  Similar as for the \ACode{\ACon{Traversal}}, we derive further
  definitions from the axiom of a \ACode{\ACon{ComposeTraversal}}:
  \begin{itemize}
  \item 
    A map \ACode{ϕ} followed by a weakening is equivalent to a
    weakening followed by \ACode{ϕ} that has been lifted over the weakened
    variable:
    \LibCode\KCommLiftWeaken
    % \LibCode\KCommLiftWeakenProof
    By applying \ACode{\AField{⋯-fusion}} to this lemma, we get the same
    statement for multiple map applications:
    \LibCode\KCommLiftWeakenTraverse
    % \LibCode\KCommLiftWeakenTraverseProof
  \item 
    A weakening followed by a singleton substitution cancel each other out:
    \LibCode\KWeakenCancelsSingle
    % \LibCode\KWeakenCancelsSingleProof
    By applying \ACode{\AField{⋯-fusion}} and \ACode{\AField{⋯-id}} to
    this lemma, we get the same statement for multiple map applications:
    \LibCode\KWeakenCancelsSingleTraverse
    % \LibCode\KWeakenCancelsSingleTraverseProof
  \item 
    A singleton substitution followed by some map \ACode{ϕ} is the same
    as \ACode{ϕ} lifted over the variable of the singleton substitution
    followed by the singleton substitution with \ACode{ϕ} applied to
    its content:
    \LibCode\KDistLiftSingle
    % \LibCode\KDistLiftSingleProof
    By applying \ACode{\AField{⋯-fusion}} to this lemma, we get the
    same statement for multiple map applications:
    \LibCode\KDistLiftSingleTraverse
    % \LibCode\KDistLiftSingleTraverseProof
  \item 
    Given the \ACode{\ADef{⋯-↑-wk}} lemma, we can define the four
    \ACode{\ACon{ComposeKit}} instances:
    \LibCode\KComposeKitInstances
    \LibCode\KComposeKitInstancesConcrete
  \end{itemize}

  \subsection{Instantiation for System F}
  In this subsection, we show how to instantiate the
  \ACode{\ACon{ComposeTraversal}} abstraction for System F.
  In practice, this is done by our reflection algorithm automatically,
  but it can be instructive to see what happens under the hood.

>>>>

  To fill this gap, we introduce a new abstraction that we call a
  \ACode{\ACon{ComposeKit}},
  which provides an operation \ACode{\AField{\_\&/⋯\_}} that unifies
  \ACode{\ADef{\_\&\_}} and \ACode{\ADef{\_⋯\_}}. This allows us to
  define a general composition, which informally can be described as:
  \ExampleCode\KComposition

@@

  In this subsection, we define the \ACode{\ACon{ComposeKit}}
  abstraction, which provides the \ACode{\AField{\_\&/⋯\_}} operation
  that is necessary for defining composition, and the
  \ACode{\ACon{ComposeTraversal}}, which provides the
  \ACode{\AField{⋯-fusion}} lemma for general maps.

  Similarly as it was the case for \ACode{\ACon{Kit}} and \ACode{\ACon{Traversal}},
  the \ACode{\ACon{ComposeTraversal}} needs to be instantiated for
  abstract \ACode{\ACon{ComposeKit}}s, which then provides the actual
  four \ACode{\ACon{ComposeKit}} instances. This allows us to break
  the dependencies between the different \ACode{\AField{⋯-fusion}} lemmas,
  similar to how \ACode{\ACon{Kit}} and \ACode{\ACon{Traversal}} allowed
  us to break the dependency between traversals with renamings and substitutions.

  For this purpose, the \ACode{\ACon{ComposeKit}} abstraction, needs
  to satisfy two requirements:
  it needs to provide us with enough machinery to define composition, but also
  it needs to provide us with enough axioms that we are able to instanciate
  the \ACode{\ACon{ComposeTraversal}}. It is defined as followed:
  \LibCode\KComposeKit
  The \ACode{\AField{\_\&/⋯\_}} operation takes a variable or term
  (according to \ACode{K₁}) and a renaming or substitution
  (according to \ACode{K₂}) and performs the map traversal returning
  a variable or term (according to \ACode{K₁⊔K₂}).
  The third parameter \ACode{K₁⊔K₂} can be seen as a functional
  dependency and is completely determined by the choice of \ACode{K₁}
  and \ACode{K₂}.

@@

  Similar as for the \ACode{\ACon{Traversal}}, we derive further
  definitions from the axiom of a \ACode{\ACon{ComposeTraversal}}:
  \begin{itemize}
  \item 
    A map \ACode{ϕ} followed by a weakening is equivalent to a
    weakening followed by \ACode{ϕ} that has been lifted over the weakened
    variable:
    \LibCode\KCommLiftWeaken
    % \LibCode\KCommLiftWeakenProof
    By applying \ACode{\AField{⋯-fusion}} to this lemma, we get the same
    statement for multiple map applications:
    \LibCode\KCommLiftWeakenTraverse
    % \LibCode\KCommLiftWeakenTraverseProof
  \item 
    A weakening followed by a singleton substitution cancel each other out:
    \LibCode\KWeakenCancelsSingle
    % \LibCode\KWeakenCancelsSingleProof
    By applying \ACode{\AField{⋯-fusion}} and \ACode{\AField{⋯-id}} to
    this lemma, we get the same statement for multiple map applications:
    \LibCode\KWeakenCancelsSingleTraverse
    % \LibCode\KWeakenCancelsSingleTraverseProof
  \item 
    A singleton substitution followed by some map \ACode{ϕ} is the same
    as \ACode{ϕ} lifted over the variable of the singleton substitution
    followed by the singleton substitution with \ACode{ϕ} applied to
    its content:
    \LibCode\KDistLiftSingle
    % \LibCode\KDistLiftSingleProof
    By applying \ACode{\AField{⋯-fusion}} to this lemma, we get the
    same statement for multiple map applications:
    \LibCode\KDistLiftSingleTraverse
    % \LibCode\KDistLiftSingleTraverseProof
  \item 
    We define \ACode{\ACon{ComposeKit}} instances for each of the four
    combinations between renamings and substitutions:
    \LibCode\KComposeKitInstances
    Interestingly, the \ACode{\ADef{Cₛ}} instance has another
    \ACode{\ADef{ComposeKit}} parameter.
    This allows us to first use \ACode{\ADef{Cᵣ}} as an argument,
    which yields the \ACode{\ADef{ComposeKit}} between substitutions
    and renamings, and then use this resulting \ACode{\ADef{ComposeKit}}
    as an argument for the case with two substitutions.
    As shown below, Agda's instance resolution is up for the task:

    % To get the \ACode{\ADef{ComposeKit}} between substitutions and
    % renamings, the argument \ACode{\ADef{Cᵣ}} is used.
    % The resulting \ACode{\ADef{ComposeKit}} is then again used as an argument
    % for \ACode{\ADef{Cₛ}} to yield the \ACode{\ADef{ComposeKit}}
    % between two substitutions. As shown below, Agda's instance
    % resolution is up for the task:
    \LibCode\KComposeKitInstancesConcrete

  \end{itemize}

  \subsection{Instantiation for System F}
  In this subsection, we show how to instantiate the
  \ACode{\ACon{ComposeTraversal}} abstraction for System F.
  In practice, this is done by our reflection algorithm automatically,
  but it can be instructive to see, as it motivates the axioms of the
  \ACode{\ACon{ComposeKit}}.
